---
title: "HW1"
author: "Jackson Dial"
date: '2023-01-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

_Working with data_

## Q1

```{r}
rain.df <- read.table("HW1_data/rnf6080.dat.txt") 
# rain.df1 <- read.table("HW1_data/rnf6080.dat.txt", header = TRUE) #I think the header statement messes it up
```

### A

I used the _read.table()_ function to read in the data.

### B

```{r}
dim(rain.df)
```

There are 5070 rows and 27 columns. I used the _dim()_ function to find the length and width of the data frame.

### C

```{r}
colnames(rain.df)
```

To get the names of the columns, I used the _colnames()_ function, and they are named with 'V' followed by the index of the column (1:27).

### D

```{r}
rain.df[2,4]
```
I indexed the value by specifying the row number and column number inside brackets.

### E

```{r}
rain.df[2,]
```

I uses the same function as before, the brackets, but left the column argument blank as to indicate I want every column, while specifying the row.

### F

```{r}
names(rain.df) <- c("year","month","day",seq(0,23))
```

This function replaces the column names with the strings "year", "month", and "day", followed by the sequence of numbers 0:23.

