---
title: "Dial-Jackson-homework2"
author: "Jackson Dial"
date: '2023-01-15'
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Q1

## A

### Task 3

```{r}
obs_data <- rbinom(n = 100, size = 1, prob = .01)

myBernLH <- function(obs_data, theta){
  N <- length(obs_data)
  x <- sum(obs_data)
  LH <- (theta^x)* ((1-theta)^(N-x))
  return(LH)
}

#create grid
theta_sim <- seq(0, 1, length=1000)
#store the LH values
sim_lh <- myBernLH(obs_data, theta_sim)
#create the plot
head(theta_sim)
tail(sim_lh)
plot(theta_sim, sim_lh, type = "l", main = "Likelihood Profile", xlab = "Simulated Support", ylab = "Likelihood")
```

## B

### Task 4

```{r}
set.seed(1234)

obs_data <- rbinom(n = 100, size = 1, prob = .01)

BetaBernMod <- function(a,b,data){
  sum_data <- sum(data)
  n <- length(data)
  a1 <- sum_data + a
  b1 <- n - sum_data + b
  return (c(a1, b1))
}

BetaBernMod(1,1,obs_data)
BetaBernMod(3,1,obs_data)
```

## C

### Task 5

```{r}
set.seed(1234)
obs_data <- rbinom(n = 100, size = 1, prob = .01)
theta_sim <- seq(0, 1, length=1000)

n <- length(obs_data)
x <- sum(obs_data)
like <- dbeta(theta_sim, x +1, n-x+1)

#make the non informative distributions
parameters_non <- BetaBernMod(1,1,obs_data)
non_prior <- dbeta(theta_sim, 1,1)
non_post <- dbeta(theta_sim, parameters_non[1], parameters_non[2])

#make the informative distributions
parameters_info <- BetaBernMod(3,1,obs_data)
info_prior <- dbeta(theta_sim, 3,1)
info_post <- dbeta(theta_sim, parameters_info[1], parameters_info[2])

#plot all 3 with the non informative prior
plot(theta_sim, like, type = "l", col = "red", lwd = 2)
lines(theta_sim, non_prior, type = "l", lwd = 2, col = "blue")
lines(theta_sim, non_post, type = "l", lty = 2, lwd = 2, col = "green")
legend(.5, 30, legend = c("Likelihood", "Prior", "Posterior"), col = c("red", "blue", "green"), lty = c(1,1,2), lwd = c(2,2,2))

#plot them all together with the informative prior
plot(theta_sim, like, type = "l", col = "red", lwd = 2)
lines(theta_sim, info_prior, type = "l", lwd = 2, col = "blue")
lines(theta_sim, info_post, type = "l", lwd = 2, col = "green")
legend(.5, 30, legend = c("Likelihood", "Prior", "Posterior"), col = c("red", "blue", "green"), lty = c(1,1,1), lwd = c(2,2,2))
```

# Q2

## A

![](IMG-0002)

## B

The posterior is a proper pdf because it has the form of a known distribution, the Gamma distribution.

## C

```{r}
#sum of all data = 154.5
n <- 8

a <- .1
b <- 1
th <- seq(0, 1, length = 500)

x <- 154.5 #not sure if this is actually correct

like <- (th ^ n) * exp(-th * x)
prior <- dgamma(th, a, b)
post <- dgamma(th, n + a, x - b)

#plot the likelihood function
plot(
  th,
  like,
  type = "l",
  ylab = "Density",
  lty = 3,
  lwd = 3,
  xlab = expression(theta)
)

#plot the prior distribution
plot(
  th,
  prior,
  type = "l",
  ylab = "Density",
  lty = 3,
  lwd = 3,
  xlab = expression(theta)
)

#plot the posterior distribution
plot(
  th,
  post,
  type = "l",
  ylab = "Density",
  lty = 3,
  lwd = 3,
  xlab = expression(theta)
)

#plot the prior and posterior together
plot(
  th,
  prior,
  col = "blue",
  type = "l",
  ylab = "Density",
  lty = 3,
  lwd = 3,
  xlab = expression(theta)
)
lines(
  th,
  post,
  col = "red",
  type = "l",
  lty = 3,
  lwd = 3
)
legend(
  0.5,
  20,
  legend = c("Prior", "Posterior"),
  col = c("blue", "red"),
  lwd = c(3, 3),
  lty = c(3, 3)
)
```



## D

An exponential model is typically used to model the expected time until an event occurs. This event could be the time from now until the next hurricane strikes Florida. 

An exponential model would not be a good fit for something like the number of times that an individual will experience a hurricane in his/her/their lifetime.

# Q3

## A


```{r}
#plot galenshore
dgalenshore <- function(x, a, theta){
  return ((2/gamma(a)) * theta^(2 * a) * exp(-theta^2 * x^2))
}
x <- seq(0.01, 2.5, 0.01)
plot(x, dgalenshore(x, 1, 1), type = "l", ylab = "Galenshore Density(x)", col = "black", 
    ylim = c(0, 5), lwd = 2)
lines(x, dgalenshore(x, 1, 1.5), type = "l", col = "red", lwd = 2, lty = 2)
lines(x, dgalenshore(x, 3, 1), type = "l", col = "blue", lwd = 2, lty = 3)
legend("topright", c("Galenshore(1,1)", "Galenshore(1,2)", "Galenshore(3,1)"), 
    col = c("black", "red", "blue"), lwd = c(2, 2, 2), lty = c(1, 2, 3))

```

![](IMG-0003)
![](IMG-0004)
![](IMG-0005)
![](IMG-0006)
![](IMG-0007)
![](IMG-0008)
